<!---
Version: "2014-11-22"
Copyright (C) 2014 Hanna Meyer, Thomas Nauss, GNU GPL (>=3)
-->
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(sp)
library(rgdal)
library(car)

setwd("D:/active/bis-fogo/data/")
data_2007 <- readOGR("field-campaign_2007/IE_2007_pontos_esp.shp",
                     "IE_2007_pontos_esp")

#### Process data set ##########################################################
#only data with correct height information
data_2007 <- data_2007[data_2007$ALT_GPS_M>0,]

#convert from factor to numeric
data_2007$CYD_OBL <- as.numeric (data_2007$CYD_OBL)

# calculate richness
richness <- rowSums(data_2007@data[,14:89])

# split into training and test data
train_numbers <- sample(1:length(richness),length(richness)*0.75) #pick random 75% of data
elevation_research_plots <- data_2007$ALT_GPS_M[-train_numbers]
```
##### Using linear models for prediction
Once a linear regression model has been derived, it could be used for predicting
values of the dependent variable at locations where only the independent variable
is avialable (assuming that the model relationship is still valid at these
locations).

To illustrate this with the data set at hand, we artificially deleted some
of the values of the plant richness and fitted the following linear model to
the remaining value pairs of plant richness and elevation.
```{r, echo=FALSE}
#linear regression:
linear_model <- lm(richness[train_numbers] ~ data_2007$ALT_GPS_M[train_numbers])
plot(data_2007$ALT_GPS_M[train_numbers], richness[train_numbers],
     xlab="Elevation in m",ylab="Plant richness")
regLine(linear_model)
summary(linear_model)
```
As one can see, this looks pretty much the same as in the last example where we
derived the linear regression model using all available observations.

However, if no further validation of the predicted values will be performed, the
only estimate of the explanatory power is the r squared value of our linear
regression (i.e. roughly 27% explanation). Therefore we have to make sure that
at least the residuals of our linear model are normaly distributed:
```{r}
qqPlot(linear_model)
```
Obviously, that is the case (otherwise we could try to solve the problem by
transforming the input variables).

##### Computing predictions
In order to predict the plant richness for some research plots where we do not
have observational plant richness data available, we will apply the linear model
equation to these elevation values. Hence, all we need is the intercept and
slope value and there we go:
```{r}
intercept <- linear_model$coefficients[1]
slope <- linear_model$coefficients[2]
plant_richness_predicted <- slope*elevation_research_plots + intercept
```
To illustrate the result, we add the predicted plant richness to the scatter 
plot:
```{r}
plot(data_2007$ALT_GPS_M[train_numbers], richness[train_numbers],
     xlab="Elevation in m",ylab="Plant richness")
regLine(linear_model)
points(data_2007$ALT_GPS_M[-train_numbers],plant_richness_predicted,col="blue")
```

Unsurprisingly, they all fall on the regression line (i.e. the linear model).