<!---
Version: "2014-11-22"
Copyright (C) 2014 Hanna Meyer, Thomas Nauss, GNU GPL (>=3)
-->
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(sp)
library(rgdal)
library(car)

setwd("D:/active/bis-fogo/data/")
data_2007 <- readOGR("field-campaign_2007/IE_2007_pontos_esp.shp",
                     "IE_2007_pontos_esp")

#### Process data set ##########################################################
#only data with correct height information
data_2007 <- data_2007[data_2007$ALT_GPS_M>0,]

#convert from factor to numeric
data_2007$CYD_OBL <- as.numeric (data_2007$CYD_OBL)

# calculate richness
richness <- rowSums(data_2007@data[,14:89])

# split into training and test data
train_numbers <- sample(1:length(richness),length(richness)*0.75) #pick random 75% of data
```

##### Scatterplot
To visualize the dependence between two variables, a scatterplot should be the 
first choice.

The following example illustrates the dependence between plant richness and
elevation using the `plot()` function.
```{r}
plot(data_2007$ALT_GPS_M, richness,xlab="Altitude in m",ylab="Plant richness")
```
Although not a hard one, some kind of linear relationship can be observed with 
decreasing plant richness with altitude.

##### Linear model
#### Computation
To model the relationship between two linearily related variables, a simple 
linear regression can be used.

The following example illustrates the computation of a linear model between 
plant richness and elevation.
```{r}
linear_model <- lm(richness ~ data_2007$ALT_GPS_M)
print(linear_model)
```
The resulting model has an intercept of about 13.1 and a slope of about -0.01,
i.e. decreasing plant richness with elevation. To visualize that, we can 
add the regression line to the scatter plot using the `regline()` function (
in addition we add the regression equation as a legend).
```{r}
plot(data_2007$ALT_GPS_M, richness,xlab="Altitude in m",ylab="Plant richness")
regLine(linear_model)
legend("topright",legend=paste0("y = ", round(linear_model$coefficients[2],3), 
                                " * x + ",
                                round(linear_model$coefficients[1],3)),bty="n")
```

#### Goodness of fit
To estimate the goodness of fit of a linear model, one can have a look at the
coefficient of determination along with the p-value.
```{r}
summary(linear_model)
```
One can see from the summary, that the model is highly significant with a p-value
of much less than 0.01. Looking at an r squared of about 0.23, the model 
explains about 23% of the variability of the data set, i.e. one can explain 23% 
of the observed plant richness by the elevation of the observational plots.
